---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#"
)
knitr::opts_knit$set(upload.fun = knitr::imgur_upload, base.url = NULL)  

```

#rentrez

`rentrez` provides functions that work with the [NCBI Eutils](http://www.ncbi.nlm.nih.gov/books/NBK25500/)
API to search, download data from, and otherwise interact with NCBI databases.

```
library(devtools)
install_github("ropensci/rentrez")
```

#Uses the EUtils API

* Read more [EUtils](http://www.ncbi.nlm.nih.gov/books/NBK25500/
* Working with the EUtils API will often require making multiple calls using the entrez package.
* Python has similar functionality in the Biopython module; http://biopython.org/wiki/Main_Page

###A Simple Example:

**Use a DOI to return the PMID of an article using `entrez_search`** 

Use an article DOI: 
[Cancer risk reduction and reproductive concerns in female BRCA1/2 mutation carriers](http://dx.doi.org/10.1007/s10689-007-9171-7). DOI of 10.1007/s10689-007-9171-7. 

```{r}
library(rentrez)
#wcancer_paper <- entrez_search(db="pubmed", term="10.1007/s10689-007-9171-7[doi]")
wcancer_paper <- entrez_search(db="pubmed", term="10.1002/ijc.21536[doi]")

wcancer_paper$ids
```

**Link data NCBI has for the article using `entrez_link`** 

```{r}
wcancer_data <- entrez_link(db="all", id=wcancer_paper$ids, dbfrom="pubmed")
wcancer_data
```
```{r}
wcancer_data$links
```
In R the wcancer_data object cantains a


###Getting data from an article

1. We've found an article on [Cancer risk reduction and reproductive concerns in female BRCA1/2 mutation carriers](http://dx.doi.org/10.1007/s10689-007-9171-7). With a DOI of 10.1007/s10689-007-9171-7. 
2. We will find the PMID via the DOI using the entrez function `entrez_search`.

```{r doi}
library(rentrez)
wcancer_risk  <- entrez_search(db = "pubmed", term = "10.1016/j.ympev.2010.07.013[doi]")
wcancer_risk <- entrez_search(db="pubmed", term="10.1007/s10689-007-9171-7[doi]")
wcancer_risk$ids
```

Now, what sorts of data are available from other NCBI database for this paper?

```{r links}
wcancer_data <- entrez_link(db="all", id=wcancer_risk$ids, dbfrom="pubmed")
wcancer_data
```
In this case all the data is in the `links` element -- the databases linked off the PMID in NCBI:

```{r showlinks}
wcancer_data$links
```

```{r}
wcancer_proteins <- entrez_fetch(db="protein", id=wcancer_data$links$pubmed_gene, rettype="fasta")
cat(substr(wcancer_proteins, 1, 237))
```


